The Revisited Hidden Weight Bit Function

### 2024-11-26 old.tex
Please check Proposition 3 and Theorem 1

Throughout this paper, $n$ will always denote a positive integer. Further, our intervals will only contain integers, so that for instance $[0,n]=\{0,1,\dots,n\}$. The set of binary vectors of length $n$ will be denoted by $\F_2^n$. The entries of a binary vector $x\in\F_2^n$ will always be denoted by $x_1,\dots,x_n$. We denote the scalar product of two binary vectors $x,y\in\F_2^n$ by $x\cdot y=\sum_{i=1}^nx_iy_i\in\F_2$. The Hamming weight of a binary vector $x\in\F_2^n$ is denoted by $\w(x)=|\{i\in [1,n]:x_i=1\}|$.

@@

\begin{definition}[Boolean function, support]\label{def:bool_f}
	A Boolean function in $n$ variables is a function from $\F_2^n$ to $\F_2$. The set of all Boolean functions in $n$ variables is denoted by $\BN$, and we denote this set without the null function by $\BN^*$.
\end{definition}

@@

\begin{definition}[Support, balancedness]\label{def:balancedness}
	Consider a Boolean function $f\in\BN$. Then the support of $f$ is defined as $\supp(f)=\{x\in\F_2^n:f(x)\neq 0\}$. The function $f$ is called balanced if it satisfies the equalities $|\supp(f)|=2^{n-1}=|\supp(f+1)|$.

%	For $k\in [0,n]$ the function is said balanced on the slice $k$ if $||\suppk{k}(f)|-|\suppk{k}(f+1)| |\leq 1$. In particular when $|\Ekn{k}{n}|$ is even $|\suppk{k}(f)|=|\suppk{k}(f+1)|=|\Ekn{k}{n}|/2$.
\end{definition}

\begin{definition}[Algebraic Normal Form (ANF) and degree]\label{def:anf}
	We call Algebraic Normal Form of a Boolean function $f\in\BN$ its unique representation as an element of the ring $\F_2[x_1,\dots,x_n]/(x_1^2+x_1,\dots,x_n^2+x_n)$, and we represent it as follows:
	\[f(x_1,\dots,x_n)= \sum_{I \subseteq [1,n]} a_I \left( \prod_{i \in I} x_i \right), \]%x_1,\dots,x_n=\sum_{I \subseteq [1,n]} a_I x^I,\]
	where $a_I\in \F_2$. The (algebraic) degree of $f$ is defined by $\degg(f)=0$ in case $f$ is the null function, and $\degg(f)=\max\{ |I|:I\subseteq[1,n],a_I=1\}$ otherwise.
\end{definition}

\begin{definition}[Walsh transform and restricted Walsh transform]\label{def:walsh_transform}
	The Walsh transform at $a \in \F_2^n$ of a Boolean function $f\in\BN$ restricted to a subset $S\subseteq \F_2^n$ is defined as:
    \[  \wt{f,S} (a) = \sum_{x\in S} (-1)^{f(x)+a \cdot x}.\]
    The Walsh transform of $f$ is then defined as $\wt{f}=\wt{f,\F_2^n}$. For $k\in[0,n]$, we also set $\wt{f,k}=\wt{f,\Ekn{k}{n}}$.
\end{definition}

@@

\begin{definition}[Nonlinearity \eg{}~\cite{Carlet20} Page 79]%and weightwise nonlinearity
	\label{def:nl}
	The nonlinearity $\NL(f)$ of a Boolean function $f\in \BN$ is the minimum Hamming distance between $f$ and all the affine functions in $\BN$:
	\[ \NL(f) = \min\{ \hd(f,g):g\in\BN,\degg(g)\leq 1\}{,} \]
	where the Hamming distance between $f$ and $g$ is defined as $\hd(f,g)=|\{x\in\F_2^n:f(x)\neq g(x)\}|$. Alternatively, the nonlinearity of $f\in\BN$ can also be defined in terms of its Walsh transform:
	\[ \NL(f) = 2^{n-1}- \frac{1}{2} \max_{a\in \F_2^n}{|\wt{f}(a)|}. \]
\end{definition}

@@

\begin{definition}[Weightwise degree-\textit{d} functions, \cite{DAM:MeaOza24} Definition 12]\label{def:wwdegd}
    A Boolean function $f\in\BN^*$ is called weightwise degree-$d$ if it can be written under the form $f=\sum_{k=0}^n f_k \phikn{k}{n}$ with $f_k\in\BN$ of degree at most $d$. The set of weightwise degree-$d$ functions is denoted by $\WD{d}{n}$.
	Additionally, a weightwise degree-$d$ function $f=\sum_{k=0}^n f_k \phikn{k}{n}$ is called a cyclic weightwise degree-$d$ function if for all $k\in [0,n]$ and all $x\in\F_2^n$, it holds that $f_k(x)=f_0(O^k(x))$, where $O^k:\F_2^n\to\F_2^n$ is the cyclic shift by $k$ positions, defined by $O^k(x_1,\dots,x_n)=(x_{(1+k)\bmod n},\dots,x_{(n+k)\bmod n})$, the representatives modulo $n$ being taken as integers in $[1,n]$.
\end{definition}

@@

\begin{definition}[Hidden Weight Bit Function (\hwbf{})]\label{def:hwbf}
    We call Hidden Weight Bit Function (\hwbf{}) the Boolean function $h\in\BN$ defined by:
	\[h(x)= \left\{ 
	\begin{array}{l l}
	0 & \quad \text{ if } \w(x) =0, \\
	x_k & \quad \text{ if } \w(x)=k.
	\end{array} \right. \]

@@

\begin{Prop}[Properties of some functions from $\WD{1}{n}$\label{prop:wwd1}]
	Let $n\geq 2$ be an even integer.
	\begin{itemize}
		\item (\eg{}~\cite{DCC:DalMaiSar06}, Theorem $3$) The majority function $\mathsf{Maj}_n$ introduced in Definition~\ref{def:maj} satisfies:
		\[
		\NL(\mathsf{Maj}_n)=2^{n-1}-\binom{n-1}{\frac n2}.
		\]
		\item (\cite{DAM:WCST14}, Theorem $3$) The \hwbf{} $h$ introduced in Definition~\ref{def:hwbf} satisfies:
		\[
		\NL(h)=2^{n-1}-2\binom{n-2}{\frac{n-2}{2}}.
		\]	
		\item (\cite{DAM:WCST14}, Theorem $1$) The \hwbf{} $h$ introduced in Definition~\ref{def:hwbf} satisfies $\degg(h)=n-1$ for all $n\geq 4$.
	\end{itemize}
\end{Prop}

\subsection{Krawtchouk polynomials}
We use Krawtchouk polynomials and some of their properties to prove one of our main results. We give the necessary preliminaries here and refer to \eg{}~\cite{book:MacSlo78} for more details.

@@

\begin{definition}[Revisited \hwbf{}]\label{def:revHWBF}
Let $n\in \N^*$, even, we call revisited \hwbf{} and denote by $f$ the Boolean function defined as:
\[f(x_1,\cdots,x_n)=\left(\sum_{i=1}^{n/2} (x_i+1) x_{i+\frac{n}{2}}\right) + \sum_{k=1}^n \phikn{k}{n} x_k.\]
\end{definition}

@@

Since $f$ is the sum of a quadratic function and a weightwise affine function, it is a weightwise quadratic function. 
Using the formalism of Definition~\ref{def:wwdegd}, $f$ is the weightwise quadratic function defined by:% $f_0=0$ and for $k\in [1,n]$ $f_k=x_k + \sum_{i=1}^{n/2} (x_i+1) x_{i+n/2}$.
\[f_0=0, \quad \text{ and } \quad \forall k \in [1,n], \, f_k=x_k+\sum_{i=1}^{n/2} (x_i+1) x_{i+\frac{n}{2}}.\]

This form allows to derive the restricted Walsh transform of $f$, which can be a useful tool to study the balancedness and bound the nonlinearity of a function.
%Maybe add examples, such as OM24, M24

%We study the restricted Walsh transform  of $f$ to bound its nonlinearity.
\begin{proposition}\label{prop:restrWT}
Let $n \in \N^*$ even, and $f$ be the function from Definition~\ref{def:revHWBF}, for all $k\in [0,n]$ its restricted Walsh transform in $a\in \F_2^n$ is given by:
\[\wtk{f}{k}(a)=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+\frac{n}{2}}+ c \cdot x},\]
where $c$ is the $n$-length binary vector such that $c=a+e_k+\sum_{i=n/2+1}^n e_i$ where the $e_i$ denotes the canonical vector with $1$ in position $i$.
\end{proposition}
\begin{proof}
%\begin{align*}
%	\wtk{f}{k}(a)&=\sum_{x \in \Ekn{k}{n}} (-1)^{f(x)+a \cdot x}
%	=\sum_{x \in \Ekn{k}{n}} (-1)^{x_k + \sum_{i=1}^{n/2} (x_i+1) x_{i+\frac{n}{2}}+ a \cdot x}\\
%	&=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+\frac{n}{2}}+ c \cdot x},
%\end{align*}
%\begin{align*}
\[	\wtk{f}{k}(a)=\sum_{x \in \Ekn{k}{n}} (-1)^{f(x)+a \cdot x}
	=\sum_{x \in \Ekn{k}{n}} (-1)^{x_k + \sum_{i=1}^{n/2} (x_i+1) x_{i+\frac{n}{2}}+ a \cdot x}\\
	=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+\frac{n}{2}}+ c \cdot x},\]
where $c$ denotes the $n$-length binary vector such that $c=a+e_k+\sum_{i=n/2+1}^n e_i$ and $e_i$ denotes the $i$-th canonical vector.
\end{proof}

@@

Accordingly, we can study the restricted Walsh transform of this function by analyzing the one of $\sum_{i=1}^{n/2} x_i x_{i+\frac{n}{2}}$, for all $a$. We introduce the following notations.

\begin{definition}
Let $n\in \N$ be an even number, and $a\in \F_2^n$. We denote $d_n$ the function given by: $d_n(x)=\sum_{i \text{ odd}}^n x_i x_{i+1}$, 
and we denote by $\Dnka{n}{k}{a}$ the quantity:
\[\Dnka{n}{k}{a}= \sum_{x\in \Ekn{k}{n}} (-1)^{d_n(x) +a x}\]
\end{definition}

We can give a recursive relation for $\Dnka{n}{k}{a}$.

\begin{proposition}\label{prop:recursiveDnka}
Let $n\in \N$ be an even number, $k$ an integer in the range $[0,n]$ and $a\in \F_2^n$. the Following holds on $\Dnka{n}{k}{a}$:
\begin{itemize}
	\item For $n=0$, $\Dnka{0}{k}{a}=1$ if $k=0$ and $\Dnka 0ka=0$ otherwise.
	\item if $n\geq 2$, denoting  $b=(a_1,\cdots,a_{n-2})$ and $y=(x_1,\cdots,x_{n-2})$, we get:
	\[\Dnka{n}{k}{a}=\Dnka{n-2}{k}{b}+\left((-1)^{a_{n-1}} + (-1)^{a_{n}}\right)\Dnka{n-2}{k-1}{b} + (-1)^{1+a_{n-1}+a_n} \Dnka{n-2}{k-2}{b}.\]
\end{itemize}

\begin{proof}
For $n\geq 2$ we can rewrite $\Dnka{n}{k}{a}$:

\begin{align*}
\Dnka{n}{k}{a}&=\sum_{x\in \Ekn{k}{n}} (-1)^{d_n(x) +a x}\\
&=\sum_{x\in \Ekn{k}{n-2}} (-1)^{d_{n-2}(y) +b y}+
\sum_{x\in \Ekn{k-1}{n-2}} (-1)^{d_{n-2}(y) +b y + a_{n-1}}\\
&+\sum_{x\in \Ekn{k-1}{n-2}} (-1)^{d_{n-2}(y) +b y + a_{n}}
+\sum_{x\in \Ekn{k-2}{n-2}} (-1)^{d_{n-2}(y) +b y +1 + a_{n-1}+ a_{n}}\\
&=\Dnka{n-2}{k}{b}+\left((-1)^{a_{n-1}} + (-1)^{a_{n}}\right)\Dnka{n-2}{k-1}{b} + (-1)^{1+a_{n-1}+a_n} \Dnka{n-2}{k-2}{b}.
\end{align*}

The limit cases $\Dnka 0ka=1$ if $k=0$ and $\Dnka 0ka=0$ otherwise are obtained from the values of $\Dnka{2}{k}{a}$ given in Table~\ref{tab:Dnka}.
\end{proof}

@@

We remark that it gives three different cases depending on the values of the two last elements of $a$:%Then it gives three cases:
\begin{itemize}
	\item if $a_{n-1}=a_n=0$, then $\Dnka{n}{k}{a}=\Dnka{n}{k}{b}+2\Dnka{n}{k-1}{b}-\Dnka{n}{k-2}{b}$,
	\item if $a_{n-1} \ne a_n$, then $\Dnka{n}{k}{a}=\Dnka{n}{k}{b}+\Dnka{n}{k-2}{b}$,
	\item if $a_{n-1}=a_n=1$, then $\Dnka{n}{k}{a}=\Dnka{n}{k}{b}-2\Dnka{n}{k-1}{b}-\Dnka{n}{k-2}{b}$.
\end{itemize}

@@

The recursive formula of Proposition~\ref{prop:recursiveDnka} gives different cases depending on the values of $a_{n-1}$ and $a_n$, however the same reasoning applies with any $a_i, a_{i+1}$ where $i$ is odd. 
Thereafter, the value of $\Dnka{n}{k}{a}$ depends only on the number of pairs of $a$ being $(0,0)$, $(1,1)$ or $(0,1)$ or $(1,0)$, and the values $\Dnka{2}{k}{a}$ for $k \in [0,2]$ and $a \in \{ (0,0),(1,1),(0,1)\}$. 
We give the first values of $\Dnka{2}{k}{a}$ in Table~\ref{tab:Dnka},  these values and Proposition~\ref{prop:recursiveDnka} are sufficient to determine any $\Dnka{n}{k}{a}$. 
In the following, we show how the Walsh transform of $f$ can be written in terms of $\Dnka{n}{k}{a}$. Then, in Section~\ref{sec:balancedness} we use Proposition~\ref{prop:recursiveDnka} to determine the balancedness of $f$, and in Section~\ref{sec:genfunc} we study the value of $\Dnka{n}{k}{a}$ using trough generating functions.

%The first values of $\Dnka{2}{k}{a}$ are listed in Table~\ref{tab:Dnka}

\begin{table}[H]
	\centering
	\begin{tabular}{|c|c|c|c|}
		\hline
		$a$ & $(0,0)$ & $(0,1)$ & $(1,1)$ \\ \hline
		$\Dnka{2}{0}{a}$ &$1$&$1$ & $1$\\ \hline
		$\Dnka{2}{1}{a}$ &$2$& $0$ & $-2$\\ \hline
		$\Dnka{2}{2}{a}$ &$-1$&$1$ & $-1$\\ \hline
	\end{tabular}
	\tablecaption{Values of $\Dnka{2}{k}{a}$}\label{tab:Dnka}
\end{table}

Then, we can express the restricted Walsh transform of $f$ in terms of $\Dnka{n}{k}{a}$:

\begin{proposition}[Walsh transform of $f$]\label{prop:WT}
	Let $n \in \N^*$ even, and $f$ be the function from Definition~\ref{def:revHWBF}, for all $a\in \F_2^n$ its Walsh transform is given by:
	\[ \wt{f}(a)=1+\sum_{k=1}^n \Dnka{n}{k}{b+\pi(e_k)},\]
where $\pi$ is the permutation sending the first $n/2$ elements to the odd positions, and the $n/2$ last ones to the even positions, and $b=\pi(a)+(0,1,0,1,\ldots,0,1)$.
\end{proposition}

\begin{proof}
Using Proposition~\ref{prop:restrWT}, the restricted Walsh transform of $f$ can then be written as \[
\wtk{f}{k}(a)=\sum_{x \in \Ekn{k}{n}} (-1)^{f(x)+ax}=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+\frac{n}{2}}+ c \cdot x}=\Dnka{n}{k}{d},\]
where $d$ is the permutation of $c$ obtained by sending the first $n/2$ elements to the odd positions, and the $n/2$ last ones to the even positions.

Calling $\pi$ this permutation, since $c=a+e_k+\sum_{i=n/2+1}^n e_i$, we can write $d=\pi(c)=\pi(a)+ \pi(e_k)+ (0,1,0,1,\ldots,0,1)$. 
Accordingly, if $k\leq n/2$, the vector $(\pi(e_k)+ (0,1,0,1,\ldots,0,1))$ has weight $n/2+1$, with one couple of consecutive elements (the first one with an odd index) being $(1,1)$ and all the others $(0,1)$. 
Otherwise,  when $k> n/2$, the vector $(\pi(e_k)+ (0,1,0,1,\ldots,0,1))$ has weight $n/2-1$, with one couple of consecutive elements (the first one with an odd index) being $(0,0)$ and all the others $(0,1)$.

Denoting by $b$ the vector $\pi(a)+(0,1,0,1,\ldots,0,1)$, we can write the Walsh transform of $f$ in $a$ as:
\[ \wt{f}(a)=1+\sum_{k=1}^n\Dnka{n}{k}{b+\pi(e_k)}.\qedhere\]
\end{proof}

@@

In this part we determine for which values of $n$ the function $f$ is balanced, using the expression of its Walsh transform in terms of $\Dnka{n}{k}{a}$ and the recursive relation of these quantities.

\begin{theorem}[Balancedness of $f$]\label{th:bal}
	Let $n \in \N^*$ even, and $f$ be the $n$-variable function from Definition~\ref{def:revHWBF}, the following holds on its Walsh transform in $0_n$:
	\[\wt{f}(0_n)= \left \{
	\begin{array}{l l}
	0 & \text{ if } n/2 \text{ is even, } \\
	-2 \binom{n/2-1}{(n-2)/4} & \text{ if }  n/2 \text{ is odd.}
	\end{array}\right.\]

	Accordingly, $f$ is balanced when $n \equiv 0\mod 4$ and unbalanced when $n\equiv 2 \mod 4$.

\end{theorem}

@@

\subsection{Study of $\Dnka nka$ through Generating Functions}\label{sec:genfunc}

>>>>

Throughout this paper, $n\geq 0$ will always denote a non-negative integer. Further, our intervals will only contain integers, so that for instance $[0,n]=\{0,1,\dots,n\}$. The set of binary vectors of length $n$ will be denoted by $\F_2^n$, with zero vector written as $0_n$. The canonical basis of $\F_2^n$ will be written as $\{e_1,\dots,e_n\}$; therefore, $e_i\in\F_2^n$ is the vector which is zero everywhere, except at position $i$. The entries of a binary vector $x\in\F_2^n$ will always be denoted by $x_1,\dots,x_n$. We write the scalar product of two binary vectors $x,y\in\F_2^n$ as $x\cdot y=\sum_{i=1}^nx_iy_i\in\F_2$. The Hamming weight of a binary vector $x\in\F_2^n$ is denoted by $\w(x)=|\{i\in [1,n]:x_i=1\}|$.

@@

\begin{definition}[Boolean function]\label{def:bool_f}
	A Boolean function in $n$ variables is a function from $\F_2^n$ to $\F_2$. The set of all Boolean functions in $n$ variables is denoted by $\BN$, and we denote this set without the null function by $\BN^*$.
\end{definition}

@@

\begin{definition}[Algebraic Normal Form, degree]\label{def:anf}
	We call Algebraic Normal Form of a Boolean function $f\in\BN$ its unique representation as an element of the ring $\F_2[x_1,\dots,x_n]/(x_1^2+x_1,\dots,x_n^2+x_n)$, and we represent it as follows:
	\[f(x_1,\dots,x_n)= \sum_{I \subseteq [1,n]} a_I \left( \prod_{i \in I} x_i \right), \]%x_1,\dots,x_n=\sum_{I \subseteq [1,n]} a_I x^I,\]
	where $a_I\in \F_2$. The (algebraic) degree of $f$ is defined by $\degg(f)=0$ in case $f$ is the null function, and $\degg(f)=\max\{ |I|:I\subseteq[1,n],a_I=1\}$ otherwise.
\end{definition}

\begin{definition}[Walsh transform, restricted Walsh transform]\label{def:walsh_transform}
	The Walsh transform at $a \in \F_2^n$ of a Boolean function $f\in\BN$ restricted to a subset $S\subseteq \F_2^n$ is defined as:
    \[  \wt{f,S} (a) = \sum_{x\in S} (-1)^{f(x)+a \cdot x}.\]
    The Walsh transform of $f$ is then defined as $\wt{f}=\wt{f,\F_2^n}$. For $k\in[0,n]$, we also set $\wt{f,k}=\wt{f,\Ekn{k}{n}}$.
\end{definition}

@@

\begin{definition}[Balancedness]
    A Boolean function $f\in\BN$ is called balanced if $\wt f(0_n)=0$.
\end{definition}

\begin{definition}[Nonlinearity, \eg{} page 79 in \cite{Carlet20}]%and weightwise nonlinearity
	\label{def:nl}
	The nonlinearity $\NL(f)$ of a Boolean function $f\in \BN$ is the minimum Hamming distance between $f$ and all the affine functions in $\BN$:
	\[ \NL(f) = \min\{ \hd(f,g):g\in\BN,\degg(g)\leq 1\}{,} \]
	where the Hamming distance between $f$ and $g$ is defined as $\hd(f,g)=|\{x\in\F_2^n:f(x)\neq g(x)\}|$. Alternatively, the nonlinearity of $f\in\BN$ can also be defined in terms of its Walsh transform:
	\[ \NL(f) = 2^{n-1}- \frac{1}{2} \max_{a\in \F_2^n}{|\wt{f}(a)|}. \]
\end{definition}

@@

\begin{definition}[Weightwise degree-\textit{d} functions, Definition 12 in \cite{DAM:MeaOza24}]\label{def:wwdegd}
    A Boolean function $f\in\BN^*$ is called weightwise degree-$d$ if it can be written under the form $f=\sum_{k=0}^n f_k \phikn{k}{n}$ with $f_k\in\BN$ of degree at most $d$. The set of weightwise degree-$d$ functions is denoted by $\WD{d}{n}$.
	Additionally, a weightwise degree-$d$ function $f=\sum_{k=0}^n f_k \phikn{k}{n}$ is called a cyclic weightwise degree-$d$ function if for all $k\in [0,n]$ and all $x\in\F_2^n$, it holds that $f_k(x)=f_0(O^k(x))$, where $O^k:\F_2^n\to\F_2^n$ is the cyclic shift by $k$ positions, defined by $O^k(x_1,\dots,x_n)=(x_{(1+k)\bmod n},\dots,x_{(n+k)\bmod n})$, the representatives modulo $n$ being taken as integers in $[1,n]$.
\end{definition}

@@

\begin{definition}[Hidden Weight Bit Function]\label{def:hwbf}
    We call Hidden Weight Bit Function (\hwbf{}) the Boolean function $h\in\BN$ defined by:
	\[h(x)= \left\{ 
	\begin{array}{l l}
	0 & \quad \text{ if } \w(x) =0, \\
	x_k & \quad \text{ if } \w(x)=k.
	\end{array} \right. \]

@@

\begin{Prop}\label{prop:wwd1}
	Let $n\geq 2$ be an even integer.
	\begin{itemize}
		\item (\eg{}~\cite{DCC:DalMaiSar06}, Theorem $3$) The majority function $\mathsf{Maj}_n$ introduced in Definition~\ref{def:maj} satisfies:
		\[
		\NL(\mathsf{Maj}_n)=2^{n-1}-\binom{n-1}{\frac n2}.
		\]
		\item (\cite{DAM:WCST14}, Theorem $3$) The \hwbf{} $h$ introduced in Definition~\ref{def:hwbf} satisfies:
		\[
		\NL(h)=2^{n-1}-2\binom{n-2}{\frac{n-2}{2}}.
		\]	
		\item (\cite{DAM:WCST14}, Theorem $1$) The \hwbf{} $h$ introduced in Definition~\ref{def:hwbf} satisfies $\degg(h)=n-1$ for all $n\geq 4$.
	\end{itemize}
\end{Prop}

\subsection{Krawtchouk polynomials}
We use Krawtchouk polynomials and some of their properties to prove one of our main results. We give the necessary definition here and refer to \eg{}~\cite{book:MacSlo78} for more details.

@@

\begin{definition}[Revisited \hwbf{}]\label{def:revHWBF}
For an even integer $n\geq 0$, we call revisited \hwbf{} the Boolean function $f\in\BN$ defined as:
\[f(x)=\sum_{i=1}^{n/2}(x_{2i-1}+1)x_{2i}+ \sum_{k=1}^n x_k\phikn{k}{n}(x).\]
\end{definition}

@@

Since $f$ is the sum of a quadratic function and a weightwise affine function, it is a weightwise quadratic function. 
Using the formalism of Definition~\ref{def:wwdegd}, $f$ is the weightwise quadratic function defined by $f_0(x)=0$ and $f_k(x)=\sum_{i=1}^{n/2}(x_{2i-1}+1)x_{2i}+x_k$ for all $k\in[1,n]$. This form allows to derive the restricted Walsh transform of $f$, which can be a useful tool to study the balancedness and bound the nonlinearity of a function.
%Maybe add examples, such as OM24, M24

\begin{definition}
    For an even integer $n\geq 0$, we introduce the Boolean function $d_n\in\BN$ given by:
    \[
        d_n(x)=\sum_{i=1}^{n/2}x_{2i-1}x_{2i}.
    \]
\end{definition}

%We study the restricted Walsh transform  of $f$ to bound its nonlinearity.
\begin{proposition}\label{prop:restrWT}
Let $n\geq 2$ be an even integer. Let $a\in\F_2^n$ and let $k\in[1,n]$. If we put $c=a+e_k+\sum_{i=1}^{n/2}e_{2i}$, then the restricted Walsh transform of the revisited \hwbf{} $f\in\BN$ satisfies:
\[
    \wtk{f}{k}(a)=\sum_{x \in \Ekn{k}{n}} (-1)^{d_n(x)+c\cdot x}=\wtk{d_n}{k}(c).
\]
\end{proposition}

\begin{proof}
%\begin{align*}
%	\wtk{f}{k}(a)&=\sum_{x \in \Ekn{k}{n}} (-1)^{f(x)+a \cdot x}
%	=\sum_{x \in \Ekn{k}{n}} (-1)^{x_k + \sum_{i=1}^{n/2} (x_i+1) x_{i+\frac{n}{2}}+ a \cdot x}\\
%	&=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2} x_i x_{i+\frac{n}{2}}+ c \cdot x},
%\end{align*}
%\begin{align*}
We compute:
\begin{align*}
    \wtk{f}{k}(a)&=\sum_{x \in \Ekn{k}{n}} (-1)^{f(x)+a \cdot x}\\
    &=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2}(x_{2i-1}+1)x_{2i}+x_k+a \cdot x}\\
    &=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2}x_{2i-1}x_{2i}+\left(x_k+\sum_{i=1}^{n/2}x_{2i}+a \cdot x\right)}\\
    &=\sum_{x \in \Ekn{k}{n}} (-1)^{\sum_{i=1}^{n/2}x_{2i-1}x_{2i}+c\cdot x}.\qedhere
\end{align*}
\end{proof}

@@

Accordingly, we can study the restricted Walsh transform of the \hwbf{} by analyzing the one of $d_n$. We introduce the following notations:

\begin{definition}
Let $n\geq 0$ be an even integer and $a\in\F_2^n$ a vector. Then for every integer $k$, we define $\Dnka{n}{k}{a}=0$ if $n\not\in[0,n]$, and otherwise:
\[
    \Dnka{n}{k}{a}= \sum_{x\in \Ekn{k}{n}} (-1)^{d_n(x) +a\cdot x}.
\]
\end{definition}

The $\Dnka{n}{k}{a}$ satisfy a recursive relation:

\begin{proposition}\label{prop:recursiveDnka}
Let $n\geq 0$ be an even integer. Then for all $k\in[0,n]$ and $a\in\F_2^n$, the following hold true:
\begin{itemize}
    \item We have $\Dnka 00a=1$ and $\Dnka 0ka=0$ if $k\neq 0$.
    \item If $n\geq 2$, then for $b=(a_1,\dots,a_{n-2})$, we have:
    \[\Dnka{n}{k}{a}=\Dnka{n-2}{k}{b}+\left((-1)^{a_{n-1}} + (-1)^{a_{n}}\right)\Dnka{n-2}{k-1}{b} + (-1)^{1+a_{n-1}+a_n} \Dnka{n-2}{k-2}{b}.\]
\end{itemize}
\end{proposition}

\begin{proof}
    We have that $\Dnka 00a=(-1)^0=0$ since $\Ekn 00=\{\epsilon\}$ for the empty vector $\epsilon\in\F_2^0$. That $\Dnka 0ka=0$ if $k\neq 0$ follows directly from the definition. Assume now that $n\geq 2$. By considering binary vectors $x\in\F_2^n$ as $x=(y,x_{n-1},x_n)$ for $y\in\F_2^{n-2}$, we get:
    \begin{align*}
    \Dnka{n}{k}{a}&=\sum_{x\in \Ekn{k}{n}} (-1)^{d_n(x) +a x}\\
    &=\sum_{y\in \Ekn{k}{n-2}} (-1)^{d_{n-2}(y) +b y}+
    \sum_{y\in \Ekn{k-1}{n-2}} (-1)^{d_{n-2}(y) +b y + a_{n-1}}\\
    &\qquad+\sum_{y\in \Ekn{k-1}{n-2}} (-1)^{d_{n-2}(y) +b y + a_{n}}
    +\sum_{y\in \Ekn{k-2}{n-2}} (-1)^{d_{n-2}(y) +b y +1 + a_{n-1}+ a_{n}}\\
    &=\Dnka{n-2}{k}{b}+\left((-1)^{a_{n-1}} + (-1)^{a_{n}}\right)\Dnka{n-2}{k-1}{b} + (-1)^{1+a_{n-1}+a_n} \Dnka{n-2}{k-2}{b}.\qedhere
    \end{align*}
\end{proof}

@@

\begin{remark}
    We note that Proposition~\ref{prop:recursiveDnka} gives three different cases depending on the values of the two last elements of $a$:%Then it gives three cases:
\begin{itemize}
	\item if $a_{n-1}=0=a_n$, then $\Dnka{n}{k}{a}=\Dnka{n}{k}{b}+2\Dnka{n}{k-1}{b}-\Dnka{n}{k-2}{b}$,
	\item if $a_{n-1} \ne a_n$, then $\Dnka{n}{k}{a}=\Dnka{n}{k}{b}+\Dnka{n}{k-2}{b}$,
	\item if $a_{n-1}=1=a_n$, then $\Dnka{n}{k}{a}=\Dnka{n}{k}{b}-2\Dnka{n}{k-1}{b}-\Dnka{n}{k-2}{b}$.
\end{itemize}
\end{remark}

@@

The recursive formula of Proposition~\ref{prop:recursiveDnka} gives different cases depending on the values of the pair $(a_{n-1},a_n)$. However, the same reasoning applies with any pair of the form $(a_{2i-1}, a_{2i})$ for $i\in[0,n/2]$. Therefore, the value of $\Dnka{n}{k}{a}$ depends only on the number of pairs $(a_{2i-1}, a_{2i})$ in $a$ being $(0,0)$, $(1,1)$, $(0,1)$ or $(1,0)$.
We give the values of $\Dnka{2}{k}{a}$ for $n=2$ in Table~\ref{tab:Dnka}. These values together with Proposition~\ref{prop:recursiveDnka} are sufficient to determine any $\Dnka{n}{k}{a}$.

\begin{table}[H]
	\centering
	\begin{tabular}{|c|c|c|c|}
		\hline
		$a$ & $(0,0)$ & $(0,1)$ & $(1,1)$ \\ \hline
		$\Dnka{2}{0}{a}$ &$1$&$1$ & $1$\\ \hline
		$\Dnka{2}{1}{a}$ &$2$& $0$ & $-2$\\ \hline
		$\Dnka{2}{2}{a}$ &$-1$&$1$ & $-1$\\ \hline
	\end{tabular}
	\tablecaption{Values of $\Dnka{2}{k}{a}$ for $n=2$.}\label{tab:Dnka}
\end{table}

In the following, we show how the Walsh transform of $f$ can be written in terms of the $\Dnka{n}{k}{a}$. Then, in Section~\ref{sec:balancedness} we use Proposition~\ref{prop:recursiveDnka} to determine the balancedness of $f$, and in Section~\ref{sec:genfunc} we study the value of $\Dnka{n}{k}{a}$ using generating functions.

\begin{proposition}\label{prop:WT}
    Let $n\geq 0$ be even, and let $a\in\F_2^n$. Put $b=a+\sum_{i=1}^{n/2}e_{2i}$. Then the Walsh transform of the revisited \hwbf{} $f\in\BN$ at $a$ is given by:
    \[ \wt{f}(a)=1+\sum_{k=1}^n \Dnka{n}{k}{b+e_k}.\]
\end{proposition}

\begin{proof}
For every $k\in[1,n]$, define $c_k=a+e_k+\sum_{i=1}^{n/2}e_{2i}=b+e_k$. Then, using Proposition~\ref{prop:restrWT}, we get:
\[
    \wt{f}(a)=1+\sum_{k=1}^n\wtk fk(a)=1+\sum_{k=1}^n\wtk{d_n}k(c_k)=1+\sum_{k=1}^n\Dnka nk{c_k}=1+\sum_{k=1}^n\Dnka nk{b+e_k}.\qedhere
\]
\end{proof}

@@

In this part, we determine for which values of $n$ the revisited \hwbf{} $f\in\BN$ is balanced by using the expression of its Walsh transform in terms of the $\Dnka{n}{k}{a}$ together with the recursive relation of these quantities.

\begin{theorem}\label{th:bal}
    For an even integer $n\geq 2$, the following holds true for the Walsh transform of the revisited \hwbf{} $f\in\BN$ at $0_n$:
	\[\wt{f}(0_n)= \left \{
	\begin{array}{l l}
	0 & \text{ if } n/2 \text{ is even, } \\
	-2 \binom{n/2-1}{(n-2)/4} & \text{ if }  n/2 \text{ is odd.}
	\end{array}\right.\]
	Accordingly, $f$ is balanced when $n \equiv 0\bmod 4$ and unbalanced when $n\equiv 2 \bmod 4$.
\end{theorem}

@@

\subsection{Study of $\Dnka nka$ through generating functions}\label{sec:genfunc}

\begin{definition}\label{defi:P_a}
    Let $n\geq 2$ be an even integer. For $a\in\F_2^n$, denote by $p=p(a)$ the number of odd $i\geq 1$ for which $(a_i,a_{i+1})=(0,0)$, by $q=q(a)$ the number of odd $i\geq 1$ for which $(a_i,a_{i+1})=(1,1)$, and by $r=r(a)$ the number of odd $i\geq 1$ for which $(a_i,a_{i+1})\in\{(0,1),(1,0)\}$. We then have $p+q+r=n/2$. Consider further the integer polynomial $P_a(z)$ given by the following expression:
    \[
    P_a(z)=(-z^2+2z+1)^{p}\cdot(-z^2-2z+1)^{q}\cdot(z^2+1)^{r}.
    \]
\end{definition}
